# Creating the index

```{r}
library(tidyverse)
library(globalmacrodata)
```

## Read the Data

```{r}
df_prices <- read_csv("../../data/processed/_01_Commodity_Prices.csv")
# df_trade <- read_csv("../../data/processed/_02_Trade.csv")
df_trade <- read_csv("../../data/processed/_02_Trade_total.csv")
```

## Merging

merging
```{r}
df_prices <- df_prices %>%
  # make all numeric
  mutate(across(-year, as.numeric)) %>%
  pivot_longer(-year, names_to = "commodity", values_to = "price") 
```

```{r}
df <- df_trade %>%
  left_join(df_prices, by = c("commodity", "year"))  %>%
  filter(commodity != "BARLEY") %>% # missing price data after 2021 
  select(
    country_iso3,
    year,
    commodity,
    # value = export_value_usd,
    value = trade_value_usd,
    # quantity = export_quantity_mt,
    quantity = trade_quantity_mt,
    price
  ) 
```

value = in thousands current USD
quantity = metric tons

which years have the most data = 2013

```{r}
df %>%
  group_by(year) %>%
  summarize(
    n_obs = n(),
    n_countries = n_distinct(country_iso3),
  ) %>%
  arrange(desc(n_obs))
```


## Weighting

Problem with using 1995 as base year = many countries have no net exports for that year, but start with exports later
=> use 2013 as base year (has the most observations)

```{r}
base_year <- 1995
```

Calculate weights based on export value in base year

```{r}
df_weights <- df %>%
  filter(year == base_year) %>%
  group_by(country_iso3) %>%
  reframe(
    # weight based on export value
    commodity,
    weight_ict = (value / sum(value, na.rm = TRUE))
    ) 
```

Check weights sum to 1
```{r}
df_weights %>%
  group_by(country_iso3) %>%
  summarize(
    total_weight = sum(weight_ict, na.rm = TRUE)
  ) %>%
  arrange(desc(total_weight)) 
```

## Index Calculation

```{r}
df_index_raw <- df %>%
  left_join(df_weights, by = c("country_iso3", "commodity")) %>%
  group_by(country_iso3, year) %>%
  reframe(
    index = prod(price ^ weight_ict, na.rm = T)
  ) 
```

```{r}
df_index_raw %>%
  filter(country_iso3 == "DEU") %>%
  ggplot(aes(x = year, y = index)) +
    geom_line() 
```

## GDP weighted

get GDP from globalmacrodatabase

```{r}
df_gmd <-  gmd(version = "2025_09")
```

```{r}
df_gdp <- df_gmd %>%
  select(ISO3, year, gdp = nGDP_USD) %>% #nominal GDP in USD
  filter(year == base_year)
```

```{r}
df_gdp_weight <- df %>%
  filter(year == base_year) %>%
  group_by(country_iso3) %>%
  reframe(
    gdp = df_gdp$gdp[match(country_iso3, df_gdp$ISO3)] * 1000, # convert to current USD
    commodity_weight = value / gdp,
    gdp_weight = sum(commodity_weight, na.rm = TRUE)
  ) %>%
  select(country_iso3, gdp_weight) %>%
  distinct()

df_gdp_weight %>%
  arrange(desc(gdp_weight))
```

unrealisitc countries: VEN, SLE, LBR, CHE

```{r}
df_index_gdp_weighted <- df_index_raw %>%
  left_join(df_gdp_weight, by = "country_iso3") %>%
  # filter(!country_iso3 %in% c("VEN", "SLE", "LBR")) %>%
  group_by(country_iso3, year) %>%
  reframe(
    index_gdp_weighted = gdp_weight * index 
  )
```

```{r}
df_index_gdp_weighted %>%
  filter(
    country_iso3 %in% c(
      # make the plot for 30 large economies
      "USA", "CHN", "JPN", "DEU", "IND", "GBR", "FRA", "ITA", "BRA", "CAN",
      "RUS", "KOR", "ESP", "AUS", "MEX", "IDN", "NLD", "SAU", "TUR",
      "ARG", "SWE", "POL", "BEL", "THA", "NGA", "AUT", "IRN", "NOR", "ISR"
    )
  ) %>%
  ggplot(aes(x = year, y = index_gdp_weighted, color=country_iso3)) +
    geom_line()
```

