# Combining the data and producing summary statistics

```{r}
library(tidyverse)
library(haven)
library(writexl)
```


## Load the datasets

First, the processed datasets from different sources

```{r}
df_wesp <- read_dta("../../data/processed/01_WESP.dta")
df_gmd <- read_dta("../../data/processed/02_GMD.dta")
df_wdi <- read_dta("../../data/processed/03_WDI.dta")
df_imf <- read_dta("../../data/processed/04_IMF_data.dta")
df_wid <- read_dta("../../data/processed/05_WID.dta")
df_conf <- read_dta("../../data/processed/06_Conf_Board.dta")
df_index <- read_dta("../../data/processed/07_Misc_Ind.dta")
df_wefm <- read_dta("../../data/processed/08_WEFM.dta")
```

Then, our newly created Commodity Index dataset
```{r}
df_comm <- read_dta("../../data/intermediate/03_Index.dta")
```

## Combine the datasets

Some datasets have the year variable as a character, we need to convert it to numeric before merging. 
```{r}
df_gmd <- df_gmd %>% mutate(year = as.integer(year))
df_wdi <- df_wdi %>% mutate(year = as.integer(year))
df_imf <- df_imf %>% mutate(year = as.integer(year))
df_wid <- df_wid %>% mutate(year = as.integer(year))
df_conf <- df_conf %>% mutate(year = as.integer(year))
df_comm <- df_comm %>% mutate(year = as.integer(year))
df_wefm <- df_wefm %>% mutate(year = as.integer(year))
```

Where the \* Magic \*  happens
```{r}
df_combined <- df_wefm %>%
  left_join(df_wesp, by = c("ISO3")) %>%
  full_join(df_gmd, by = c("ISO3", "year")) %>%
  full_join(df_wdi, by = c("ISO3", "year")) %>%
  full_join(df_imf, by = c("ISO3", "year")) %>%
  full_join(df_wid, by = c("ISO3", "year")) %>%
  full_join(df_conf, by = c("ISO3", "year")) %>%
  full_join(df_index, by = c("ISO3", "year")) %>%
  full_join(df_comm, by = c("ISO3", "year"))
```

Make sure that countries align with our WESP countries
```{r}
df_combined <- df_combined %>%
  filter(ISO3 %in% df_wesp$ISO3) 
```


## Save the combined dataset

```{r}
write_dta(df_combined, "../../data/processed/final_data.dta")
write_csv(df_combined, "../../data/processed/final_data.csv")
```

## Excel File

For easier analysis in the Team, an excel file with separate sheets for each of
- Productivity growth = labor_productivity_hour_growth
- Investment over GDP = inv_GDP
- R&D investments = rnd_gdp
- Researchers = researchers_per_million
- Economic complexity index = eci
- Publications = pub_per_million
- Patent Applications per Million = patents_per_million

```{r}
df_list <- list()

for (var in c("labor_productivity_hour_growth", "inv_GDP", "rnd_gdp", "eci", "researchers_per_million", "patents_per_million", "pub_per_million")) {
  df_var <- df_combined %>%
    select(ISO3, year, all_of(var)) %>%
    pivot_wider(names_from = year, values_from = all_of(var)) 
  df_list[[var]] <- df_var
}
```

```{r}
write_xlsx(
  path = "../../data/processed/innovation_data.xlsx",
  x = df_list
)
```

